Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable

### 2020-07-21 paper/flow_based.tex
adjust brownfield appendix

\section{Appendix}
\subsection{LMP from Optimization}
The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\

@@

\subsection{Example: Power Flow Allocations of different Types}
\label{sec:example_plots}
\begin{figure}[h!]
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1_net_ebe}
    \end{subfigure}
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_net_ebe}
    \end{subfigure}
    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allcated net power injections (scheme \ref{net}\ref{ebe}). Bus~1 retrieves 60~MW from itself and nothing from bus~2. The latter in turn retrieves 40~MW from bus~1 and self supplies 50~MW. The P2P trades are less in number and more intuitive then with allocating gross flow.}
    \label{fig:example_allocation_net_ebe}
\end{figure}

@@

Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of
\begin{align}
\capacityGeneration \ge \capacityGenerationLower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \\
\capacityGeneration \le \capacityGenerationUpper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
\end{align}
% or doing likewise with a subset $L$ of transmission lines,
% \begin{align}
% \capacityFlow \ge \capacityFlowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
% \capacityFlow \le \capacityFlowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
% \end{align}
alters the objective value as soon as one of those become bounding. 
The complementary slackness state is 
\begin{align}
 \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \\
 \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \\
\end{align}

The CAPEX paybacks for generators %and transmission lines
\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
\begin{align}
\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s}
\label{eq:capex_generation_duality_bf}
\end{align}
for generators. % and 
% \begin{align}
% \pdv{\lagrangian}{\capacityFlow}  = 0 \,\, \rightarrow \,\, \\
% \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
% \label{eq:capex_flow_duality_bf}
% \end{align}
% for transmission lines.
Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ 
\begin{align}
 \capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGenerationLower - \muuppergenerationnom \capacityGenerationUpper \Forall{n,s}
\end{align}
Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
\begin{align}
  \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
\end{align}

playaround

\begin{align}
\capexGeneration \, \capacityGeneration =  \sum_t \muuppergeneration \, \generation +  \mulowergenerationnom \capacityGeneration - \muuppergenerationnom \capacityGeneration \Forall{n,s}
\end{align}

>>>>

\section{Appendix}

\renewcommand\theequation{\thesection.\arabic{equation}}
\setcounter{equation}{0}

\renewcommand\thefigure{\thesection.\arabic{figure}}    
\setcounter{figure}{0}

\subsection{LMP from Optimization}
The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay.\\

@@

\subsection{Example: Power Flow Allocations of different Types}
\label{sec:example_plots}
\begin{figure}[h!]
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus1_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus1_net_ebe}
    \end{subfigure}
    \begin{subfigure}[c]{\linewidth}
    \includegraphics[width=\linewidth]{example_allocation_bus2_net_ebe.png}
    \vspace{-40pt}
    \subcaption{}
    \label{fig:example_allocation_bus2_net_ebe}
    \end{subfigure}
    \caption{Power allocations for bus~1 (a) and bus~2 (b) of the example network in \cref{fig:example_network} using equivalently allocated net power injections (scheme \ref{net}\ref{ebe}). Bus~1 retrieves 60~MW from itself and nothing from bus~2. The latter in turn retrieves 40~MW from bus~1 and self supplies 50~MW. The P2P trades are less in number and more intuitive then with allocating gross flow.}
    \label{fig:example_allocation_net_ebe}
\end{figure}

@@

Constraining the capacities $\capacityGeneration$  for a subset $S$ of generators to lower or upper limits in the form of
\begin{align}
\capacityGeneration \ge \capacityGenerationLower \resultsin{\mulowergenerationnom} \hpad \forall{n,s \in S} \\
\capacityGeneration \le \capacityGenerationUpper \resultsin{\muuppergenerationnom} \hpad \forall{n,s \in S}
\end{align}
alters the objective value as soon as one of those become bounding. 
The complementary slackness for both are 
\begin{align}
    \muuppergenerationnom \left( \capacityGeneration - \capacityGenerationUpper \right) =0 \Forall{n,s \in S} \\
    \mulowergenerationnom \left( \capacityGeneration - \capacityGenerationLower \right) =0 \Forall{n,s \in S}
\end{align}

The CAPEX paybacks for generators and transmission lines
\cref{eq:capital_price_generation_sum,eq:capital_price_flow_sum} change to 
\begin{align}
\pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, \\
\capexGeneration =  \sum_t \muuppergeneration \, \generationpotential + \mulowergenerationnom - \muuppergenerationnom \Forall{n,s \in S}
\label{eq:capex_generation_duality_bf}
\end{align}
for generators. Multiplying \cref{eq:capex_generation_duality_bf} by $ \capacityGeneration$ leads us to 
\begin{align}
 \capexGeneration \, \capacityGeneration - \totalRest_s=  \sum_t \muuppergeneration \, \generation \Forall{n,s \in S}
 \label{eq:capex_generation_duality_bf2}
\end{align}
where we define the cost resulting from the capacity expansion limits as
\begin{align}
    \totalRest_s = \left(\mulowergenerationnom - \muuppergenerationnom \right) \capacityGeneration
    \Forall{s}
\end{align}
As $\mulowergenerationnom \ge 0$ and $\muuppergenerationnom \le 0$ the latter represents a net positive cost term.

Multiplying \cref{eq:opex_duality}, which is still valid, with $\generation$ and inserting \cref{eq:capex_generation_duality_bf2} will bring us to the zero profit rule for generators with capacity expansion limits, 
\begin{align}
    \capexGeneration \capacityGeneration + \sum_t \opexGeneration \generation  - \totalRest_s = \sum_t \lmp \incidenceGenerators \generation \Forall{s \in S}
\end{align}

% or doing likewise with a subset $L$ of transmission lines,
% \begin{align}
% \capacityFlow \ge \capacityFlowLower \resultsin{\mulowerflownom} \hpad \forall{\ell \in \textit{L}} \\
% \capacityFlow \le \capacityFlowUpper \resultsin{\muupperflownom} \hpad \forall{\ell \in \textit{L}}
% \end{align}

% and 
% \begin{align}
%     \muupperflownom \left( \capacityFlow - \capacityFlowUpper \right) =0 \\
%     \mulowerflownom \left( \capacityFlow - \capacityFlowLower \right) =0 \\
% \end{align}

% \begin{align}
%     \pdv{\lagrangian}{\capacityFlow}  = 0 \,\, \rightarrow \,\, \\
%     \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right)  + \mulowerflownom - \muupperflownom \Forall{n,s}
%     \label{eq:capex_flow_duality_bf}
%     \end{align}
%     for transmission lines.

% Multiplying \cref{eq:capex_generation_duality_bf} by the amount of expansion $\left( \capacityGeneration - \capacityGenerationLower \right) $
% \begin{align}
%   \capexGeneration \, \capacityGeneration^{exp} =  \sum_t \muuppergeneration \, \left( \generation - \generationpotential \, \capacityGenerationLower \right) + \muuppergenerationnom \left( \capacityGenerationUpper - \capacityGenerationLower \right) \Forall{n,s}
% \end{align}
